MS_Imp_Guidelines

### 2023-01-26 Markdown/Manuscript/Child_RMDs/MS_Aims.Rmd
working on discussion

<!-- Add aims text here -->
In this paper, we introduce a framework for developing a modular youth mental health MOSHEM. Specifically we:

@@

<!-- We have consolidated and refined these and other recommended guidelines for OSHEMs as part of a framework for developing an open source model of youth mental health.  -->

>>>>

In this paper, we introduce a framework for developing a modular youth mental health MOSHEM. Specifically we:

@@

### 2023-01-26 Markdown/Manuscript/Child_RMDs/MS_Discussion.Rmd
stuck
safety
up to fw
working on discussion

<!-- MH systems design is not a pharma led project - less concerns about commercial ownership -->

# Motivation

## Why develop OSHEMs in mental health

Mental disorders impose high health, social and economic burdens worldwide [@RN8;@GBD2019]. Much of this burden is potentially avertable [@RN25], but poorly financed and organised mental health systems are ill-equipped for this challenge[@RN22;@RN23]. The large and widespread additional mental health burdens recently observed during the COVID-19 pandemic[@20211700] and predicted as a potential future consequence of climate change [@page_howard_2010], highlight the need to improve the resilience and adaptability of these systems. To help stem growing demand for mental health services, policymakers have also been encouraged to place greater emphasis on tackling the social determinants of mental disorders [@RN11].

Major mental health reform programs, <!-- such as those currently being implemented in Victoria, Australia [@VicRC2021],  -->can involve the identification, prioritisation, sequencing, targeting and monitoring of multiple interdependent initiatives.

@@

Developing networks of modellers working on common health conditions has been recommended as a strategy for improving model validity [@Sampson2019] and some of us are part of a nascent initiate of this type in mental health [@nmhsmn2022]. 
<!-- Add ref to NMHSMN in Australia --> Similarly, developing partnerships between modellers and decision-makers across the life-cycle of a modelling project can help ensure models are appropriately conceptualised, implemented and have practical utility as decision aids [@Zabell2021; @SQUIRES2016588].

>>>>

In this article we introduced a framework that we developed to help us implement a MOSHEM in youth mental health. We outlined the standards for accountable, reusable and updatable models specified by that framework and the modelling toolkit we created to help the youth mental health MOSHEM we are developing to meet those standards. We also provided an overview of the initial set of model modules we developed with the framework to implement a utility mapping study and summarised the extent to which the outputs of that study adheres to framework standards.

The work we have described raises a number of issues relating to the future development of a youth mental health MOSHEM, for health economists interested in undertaking other MOSHEMS and for health research funders.

## Implications for the development of a youth mental helath MOSHEM
<!-- ### Implications for our modelling team -->
Mental disorders impose high health, social and economic burdens worldwide [@RN8;@GBD2019]. Much of this burden is potentially avertable [@RN25], but poorly financed and organised mental health systems are ill-equipped for this challenge[@RN22;@RN23]. There is therefore an opportunity for economic research to help identify mental health reforms of significant potential public benefit. To date, economic studies in mental health have predominantly examined the affordability and value for money of individual programs [@RN34]. This economic evaluation work is an essential prerequisite for developing economically sustainable mental health reform plans, but it needs to be accompanied by and integrated with a broader program of economic research. The classification system developed by Wagstaff and Culyer [@wagstaff2012four] identified twelve categories of health economic research. In addition to the economic evaluation category, we have prioritised the following areas for the future development of modules for our MOSHEM:

- health and its value [utility and choice],

- demand for health and health care,

- determinants of health and ill-health,

- supply of health services,

- efficiency and equity,

In economic evaluation, 
<!-- Yet single purpose models that assume static systems may be inadequate for the decision support needs of policymakers and service planners [@PC2020; @Occhipinti2021].  -->

<!-- The large and widespread additional mental health burdens recently observed during the COVID-19 pandemic[@20211700] and predicted as a potential future consequence of climate change [@page_howard_2010], highlight the need to improve the resilience and adaptability of these systems. To help stem growing demand for mental health services, policymakers have also been encouraged to place greater emphasis on tackling the social determinants of mental disorders [@RN11].  -->

 - Aims of our model (multi-purpose, focused on systems reform)
 - Addressing wagstaff areas ()
 - Adequacy of current methods / our capacity (errors / impact)
 - One shot at big changes
 - Opportunities for others to leverage and enhance
 - Our analysis - do it in stages, be transparent, continually improve, get help, be flexible (modelling toolkit / monolithic model), have a framework

 - Strengths (comparisons to existing work, benefits to ourselves and others)

 - Limitations (technical, not behavioural / institutional solution)

<!-- A coding framework for MOSHEMs developed in the language R includes standardised approaches to directory structure and naming conventions [@Alarid2019]. -->
<!-- Our approach to model development is to undertake a number of discrete modelling projects of the people, places, platforms and programs that shape the mental health and wellbeing of young people and to progressively link them together by means of a common framework. To model people we are developing synthetic representations of populations of interest [@DVN/HJXYKQ_2021] that describe relevant individual characteristics and their relationships, algorithms that map psychological measures to health utility [@Hamilton2021.07.07.21260129] and choice models for predicting the helpseeking behaviour of young people [@hamilton_matthew_2022_6627995]. Our in-development model of places [@DVN/V3OKZV_2022] has the aim of synthesising geometry and spatial attribute data to characterise the geographic distribution of relevant demographic, environmental, epidemiological and service infrastructure features. We are in the early stages of a multi-annual project to develop a service platform model that will represent the processes and operations of a complex primary youth mental health service. We also plan to extend and update our prior work reviewing economic evidence relating to youth mental health programs [@RN33] so that it can be integrated with this model. Our initial work on **readyforwhatsnext** is focused on Victoria, Australia but the framework we are using to develop it is designed to facilitate extension by ourselves and others to different jurisdictional decision contexts. Progress is reported on a project website [@rfwn2022]. -->
Major mental health reform programs, <!-- such as those currently being implemented in Victoria, Australia [@VicRC2021],  -->can involve the identification, prioritisation, sequencing, targeting and monitoring of multiple interdependent initiatives.

@@

<!-- ### Implications for potential users of and contributors to our model -->

Another future priority is to submit production versions of each R package for review by and archiving on the Comprehensive R Archive Network (CRAN) [@CRAN2022].
Developing networks of modellers working on common health conditions has been recommended as a strategy for improving model validity [@Sampson2019] and some of us are part of a nascent initiate of this type in mental health [@nmhsmn2022].

<!-- Add ref to NMHSMN in Australia --> Similarly, developing partnerships between modellers and decision-makers across the life-cycle of a modelling project can help ensure models are appropriately conceptualised, implemented and have practical utility as decision aids [@Zabell2021; @SQUIRES2016588].

- Community building (feasibility)
<!-- Open source is more than open access (limited discussion). -->
- Funding contingent directions - replications / transfers

## Implications for health economists
<!-- Standardized and transferable approaches to health economic modelling have been recommended. [doi:10.1017/S0266462317000666] -->
<!-- As modelling methods have become more sophisticated, some have argued that more complex models are required to adequately explore certain types of public health topics [@SQUIRES2016588].  -->
<!-- Modellers should be faithful to reaity and values of users DOI: 10.1002/asi.23697 -->
<!-- transferability - [https://doi.org/10.1017/S026646232200321X].  -->
Output by health economists [@pitt2016economic; @wagstaff2012four; @munoz2020health]

## Implications for research funders
<!-- Reducing waste in research is responsibility of research funder: https://doi.org/10.1016/S0140-6736(13)62229-1 -->
<!-- MH Research priorities (WP3 - Research, WP6 - Economic) https://doi.org/10.1016/S2215-0366(15)00332-6 -->
<!-- http://hdl.handle.net/1893/27205--><!-- CONTEXT: Funders should also support methodological work on context, including the development of reporting templates to facilitate accurate, comprehensive and systematic reporting of context, approaches to evidence synthesis that can capture contextual variation adequately and modelling methods that can leverage the accumulating evidence on contextual variation to predict the reach, implementation and effectiveness of promising interventions in new contexts." -->

- unreasonable to expect more accountability without more resources

<!-- Link to Wagstaff - Demand forecasting, valuation of health and healthcare, economic evaluation, utility mapping?? -->

<!-- # Motivation -->

<!-- ## Why develop OSHEMs in mental health -->

<!-- Decision makers are frequently left to rely on economic analyses that become less valid and relevant each year. A model can grow stale when changes in the real world, whether sudden (e.g. new pandemics) or gradual (e.g. demographic), is not represented by corresponding updates to relevant model features. Even when the system being modelled remains relatively stable, new research findings may potentially invalidate aspects of a model's original conceptualisation and implementation.  -->

<!-- that rarely explores complex dynamic systems  -->
<!-- [PECUNIA] -->

<!-- Computational modelling could play an important role in developing policies to improve population mental health but this may require significant changes in the way mental health modelling projects are funded, conceptualised and implemented.  -->
 <!-- but, as with health economics more generally, OSHEMs remain rare. Greater use of open source approaches could help improve the scope, validity and usefulness of mental health economic models. -->

<!-- Repositories such as Zenodo [@Zenodo2013] and Dataverse [@Dataverse2007] provide persistent storage solutions that generate a Digital Object Identifier (DOI) for each unique item. These repositories are a preferable solution for sharing citable code and data than transitory repositories such as corporate websites or GitHub where items can be deleted or relocated at any time [@sseditors2022]. Zenodo includes tools that automate integration with GitHub, which makes it easy for developers to maintain parallel code repositories - one for disseminating the most up to date development code and the other for archiving citable code releases. --> <!-- Developers storing data in a Dataverse installation have access to multiple meta-data fields to document both a data collection and its individual constituent files.  --><!-- In R, code manuals and websites can be created with the aid of tools such as devtools [@devtools2021], sinew [@sinew2022], roxygen2 [@roxygen2021] and pkgdown [@pkgdown2022].  -->

<!-- and the use of version control systems. <!-- such as Git [@git20XX].  -->

<!-- Applying a previously published algorithm [@kross2016] to analyse the most comprehensive archive of released R packages [@CRAN2022] finds that `r gpl_pc_dbl`% are distributed under various forms of General Public License (GPL) [@GNUGPL2022], a copyleft license. -->
<!-- In R, including a CITATION file in the `inst` directory of a package will enable users of that package to retrieve citation information by running a command of the format `citation("Package Name")` in the R console [@Salmon2021]. More generally, including a CITATION.cff file at the top level of your code repository will enable GitHub and Zenodo repositories hosting that item to include the relevant information in their citation tools [@Druskat_Citation_File_Format_2021]. Datasets hosted on Dataverse installations have metadata fields that, once completed by authors, generate citation files for dataset viewers. -->

<!-- A key challenge to generalising health economic models is that they are typically developed to inform a decision problem with a highly specific jurisdictional context. However, a number of choices about how these models are implemented can significantly increase the re-usability of model code in other contexts.  -->
<!-- For example, when generalising a model developed for the Australian context to a UK context, one could create a class that initially inherits all of the methods defined for the Australian model and then write new or replacement methods as needed for the UK model.  -->

<!-- A number of tools and approaches can make the process of implementing and curating changes to model code and data more coherent and efficient.   <!-- Repositories such as Zenodo [@Zenodo2013] and Dataverse [@Dataverse2007] provide persistent access to all published versions of a dataset, each uniquely identifiable. For code, use of version control tools like Git [@git20XX] can ensure that the entire development history of a project is organised so that each version is distinguishable and retrievable by developers. The online platform GitHub [@github2007] can make this version history accessible to anyone. -->

<!-- For R code, the usethis [@usethis2021] package can be used to partially automate version number increments using the convention Major.Minor.Patch.Development (e.g. ExampleSoftware v1.1.1.1). Datasets stored on the Harvard Dataverse use the simpler Major.Minor convention.  -->
<!-- MOSHEMs developed in R can take advantage of templates provided by devtools [@devtools2021] and pkgdown [@pkgdown2022] to run continuous integration checks on GitHub.  --><!-- These tests can include those of units (do individual functions produce expected output?), documentation (does documentation render correctly?; can all example workflows be executed?) and installation (can the software be successfully deployed on multiple types of operating system?).  -->
<!-- MH systems design is not a pharma led project - less concerns about commercial ownership -->

<!-- and for mental health specifically [@Occhipinti2021],  -->
<!-- could provide greater insight about inter-dependencies between candidate policies and the dynamic nature of the mental health systems planning context. Dynamic systems methods might provide the foundations for developing   -->
<!-- and models designed for multiple purposes r -->

 <!-- as part of the Open Value Initiative [@Jansen2019].  -->

### 2023-01-26 Markdown/Manuscript/Child_RMDs/MS_Imp_Example.Rmd
working on discussion

We used framework toolkits to develop model code and datasets for implementing a utility mapping study that has previously been described [@Hamilton2021.07.07.21260129]. Table \@ref(tab:checktb) assesses that study against each TIMELY standard, 18 of which the study was able to meet. The two current standards where the study falls short are in reporting code coverage and including a user-interface. Both these items are current development priorities for our team.

>>>>

We applied the framework to develop modules for developing utility mapping models, supplying those modules with data and implementing utility mapping analyses. The study for which these tools were developed and which applied them to developing utility mapping models for a clinical mental health sample of young people has been previously described [@Hamilton2021.07.07.21260129]. The outputs we authored using the framework's modelling toolkit include:

 - development version module libraries for describing and validating youth mental health human record datasets [@hamilton_matthew_2022_6084467], scoring health utility [@hamilton_matthew_2022_6084824], specifying utility mapping models [@hamilton_matthew_2022_6116701] and implementing reproducible utility mapping studies [@gao_caroline_2022_6130155];

 - a development version library of functions for finding and using utility mapping models developed with these tools [@matthew_p_hamilton_2021_5646669];

 - data collections of synthetic populations for testing model modules [@DVN/HJXYKQ_2021] and study input and results data [@DVN/DKDIB0_2021];

 - programs for replicating all steps from data ingest to manuscript reporting [@hamilton_matthew_2022_6129906], applying utility mapping models to new data [@hamilton_matthew_2022_6416330] and generating a synthetic representation of the study dataset [@hamilton_matthew_p_2022_6321821];

 - subroutines for creating a catalogue of utility mapping models [@hamilton_matthew_2022_6116385] and generating a draft scientific manuscript [@matthew_p_hamilton_2022_5976988] for studies implemented with these modules.

When we assesses these study outputs against each framework standard, we conclude that 18 standards were wholly or mostly met (see Table \@ref(tab:checktb)). The two standards where the study outputs currently fall short are in reporting code coverage and including a user-interface. Both these items are scheduled to be addressed when we release production versions of the code libraries.

### 2023-01-26 Markdown/Manuscript/Child_RMDs/MS_Imp_Guidelines.Rmd
working on discussion

 <!-- - **Transparent**: people can easily see how a model has been implemented and tested; -->
 <!-- - **Iterative**: a model is routinely updated to maintain and improve validity; -->
 <!-- - **Modular**: models and their components can be combined to extend their scope; -->
 <!-- - **Epitomising**: model code is sufficiently general to be be re-used in multiple decision contexts; -->
 <!-- - **Licensed**: a model, its components and derivatives are persistently re-usable by other modellers; and -->
 <!-- - **Yielding**: a model can be simply, flexibly and reliably used as a decision aid. -->

@@

Model code and data should be clearly documented, potentially with different versions for technical and non-technical users [@Eddy2012]. <!-- nd --> Consistent use of meaningful naming conventions when authoring code is recommended [@Wilson_2017; @Alarid2019]. Code can be made easier to follow by using the practices of abstraction [@8717448], where only simple, high level commands are routinely exposed to reviewers, and polymorphism [@7181447], where the same command (e.g. "simulate") can be reused to implement different algorithms of the same type. Programs to implement model analyses can be made comprehensible to even non-technical users through the use of literate programming techniques <!-- that use tools like RMarkdown [@xie2018r]  -->that integrate computer code with plain English descriptions.

@@

<!-- Guidelines: -->

<!-- - **T1: Uniquely identified copies of model code and data are permanently archived in open online repositories** -->
<!-- <!-- T2 MISSING - Merged With T1 -->

<!-- - **T2: Model code and data are documented **<!-- Was T3 -->

<!-- - **T3: Model code uses a simple and consistent syntax**<!-- Was T4 -->

<!-- - **T4: Model analyses and reporting are implemented using literate programming** <!-- Was T5 -->

<!-- - **T5: Code coverage is reported** <!-- New, replaces version now moved to T4 -->

<!-- - **T6: All parts of a study analysis and reporting workflow can be reproduced and/or replicated** <!-- Was T7, Replaces Deleted T6 -->

<!-- - **L3: Model code and data are distributed with tools to support appropriate citation** <!-- New -->

<!-- and the use of version control systems. <!-- such as Git [@git20XX].  -->  
<!-- Repositories such as Zenodo [@Zenodo2013] and Dataverse [@Dataverse2007] provide persistent storage solutions that generate a Digital Object Identifier (DOI) for each unique item. These repositories are a preferable solution for sharing citable code and data than transitory repositories such as corporate websites or GitHub where items can be deleted or relocated at any time [@sseditors2022]. Zenodo includes tools that automate integration with GitHub, which makes it easy for developers to maintain parallel code repositories - one for disseminating the most up to date development code and the other for archiving citable code releases. --> <!-- Developers storing data in a Dataverse installation have access to multiple meta-data fields to document both a data collection and its individual constituent files.  --><!-- In R, code manuals and websites can be created with the aid of tools such as devtools [@devtools2021], sinew [@sinew2022], roxygen2 [@roxygen2021] and pkgdown [@pkgdown2022].  -->

### Standards for a reusable MOSHEM
<!-- ## Licensed models -->
```{r echo = F}
gpl_pc_dbl <- get_license_share("GPL")
```

@@

Storing and managing model code and data separately, can make it easier to apply models to different decision contexts and, when necessary, to selectively restrict access to data that are confidential, while disseminating all other model artefacts. Clear distinctions should be made between model modules (code that defines abstract data structures and the algorithms that can be applied to data described by these structures), model datasets (digital information such as parameter values, unit records, etc) and model analyses (code that specifies the data to be supplied to modules and selects which module algorithms to apply to that data).

The coding practice of encapsulation [@8717448] can be used to help ensure that model modules can be safely combined [@ready4oop2022]. In some cases, combining modules may mean new versions of modules have to be created that account for their effects on each other. The coding concept of inheritance [@8717448] can be used to efficiently achieve this objective as well as to facilitate selective editing of modules when transferring models to different decision contexts [@ready4oop2022]. Writing algorithms as collections of functions (short, self-contained and reusable software routines that each perform a discrete task) is recommended as good practice for scientific computing [@Wilson_2017]. Functions to implement model algorithms can be associated with module data structures (also known as a class) via a special type of function called a method. Model modules of a similar type or purpose can be efficiently distributed and documented by bundling them as code libraries. It is good practice to make available test or toy data to demonstrate the use of model algorithms [@Wilson_2017].

<!-- When bundled for distribution as libraries<!-- (for example, as R packages) -->
<!-- , functions have the potential to be widely re-used, often in contexts very different than those they were originally developed for.  --> 
<!-- A special type of function, called a method, can only be applied to a pre-defined class of data structure.  -->

<!-- ## Yielding models --><!-- MOSHEMs can be time and skills intensive for modellers to develop - but they should be easy for others to use.  -->
Statistical models are a common output of health economic evaluations, but they are often not reported in a format that enables others to confidently and reliably re-use them [@Kearns2013]. Open source approaches can help address this by disseminating code artefacts that enable easy and appropriate use of a statistical model to make predictions with new data. However, great care must be exercised when doing so if models are derived from data on human subjects as some software artefacts by default contain a copy of the source dataset. Such dataset copies must therefore be replaced (for example, with synthetic data) and the amended artefact's predictive performance then retested before any public release. Another way to make MOSHEMs easier to use is to develop simple user-interfaces for non-technical users. In the open source R language, such user-interfaces are typically developed with the Shiny package, for which a tutorial aimed at health economists is available [@SmithR2020].

<!-- Applying a previously published algorithm [@kross2016] to analyse the most comprehensive archive of released R packages [@CRAN2022] finds that `r gpl_pc_dbl`% are distributed under various forms of General Public License (GPL) [@GNUGPL2022], a copyleft license. -->
<!-- In R, including a CITATION file in the `inst` directory of a package will enable users of that package to retrieve citation information by running a command of the format `citation("Package Name")` in the R console [@Salmon2021]. More generally, including a CITATION.cff file at the top level of your code repository will enable GitHub and Zenodo repositories hosting that item to include the relevant information in their citation tools [@Druskat_Citation_File_Format_2021]. Datasets hosted on Dataverse installations have metadata fields that, once completed by authors, generate citation files for dataset viewers. -->

<!-- A key challenge to generalising health economic models is that they are typically developed to inform a decision problem with a highly specific jurisdictional context. However, a number of choices about how these models are implemented can significantly increase the re-usability of model code in other contexts.  -->
<!-- For example, when generalising a model developed for the Australian context to a UK context, one could create a class that initially inherits all of the methods defined for the Australian model and then write new or replacement methods as needed for the UK model.  -->
<!-- Guidelines: -->

<!-- - **L1: Model code is made available for re-use under copyleft or permissive licenses** -->

<!-- - **L2: Non-confidential model data is licensed for liberal re-use (subject to additional terms for de-identified human data)** -->

<!-- - **E2: Model code defines inheriting data-structures** --> <!-- Order changed-->

<!-- - **E1: Model code is distributed as libraries of classes and functions**  <!-- Modified (lost ref to OOP and merged with M2) -->

<!-- - **E3: Test data is available to demonstrate generalised applications of model code** -->

<!-- - **M1: Model code and data are stored and managed separately** <!-- Modified (lost ref to licensed and cited) -->

<!-- - **M2: Model code defines encapsulating data structures** <!--Replacement - Old version now merged with E1 -->

<!-- - **Y1: Statistical models are distributed with validated tools to support their safe and appropriate re-use** <!-- New, Replaces Old Y1 (now Y2) --> 
<!-- - **Y2: Simple user-interfaces allow non-technical users to configure and run models ** <!-- Was Y1 -->

### Standards for an updatable MOSHEM
To avoid MOSHEMs going stale - losing validity and usefulness with time - they should be routinely updated.  Each update of code and data should be uniquely identifiable and retrievable, a goal that can be facilitated by use of version control tools [@Erdemir2020]. Labeling each change using semantic versioning [@semver20xx] conventions can signal the potential importance of an update to users of model code and data. Continuous integration [@CI2017] tools can help verify that each code update passes multiple quality tests. Finally, using deprecation conventions that take an informative and staged approach to retiring old code and data reduces the risk that model revisions have unintended consequences on third party users.

<!-- A number of tools and approaches can make the process of implementing and curating changes to model code and data more coherent and efficient.   <!-- Repositories such as Zenodo [@Zenodo2013] and Dataverse [@Dataverse2007] provide persistent access to all published versions of a dataset, each uniquely identifiable. For code, use of version control tools like Git [@git20XX] can ensure that the entire development history of a project is organised so that each version is distinguishable and retrievable by developers. The online platform GitHub [@github2007] can make this version history accessible to anyone. -->

<!-- For R code, the usethis [@usethis2021] package can be used to partially automate version number increments using the convention Major.Minor.Patch.Development (e.g. ExampleSoftware v1.1.1.1). Datasets stored on the Harvard Dataverse use the simpler Major.Minor convention.  -->
<!-- MOSHEMs developed in R can take advantage of templates provided by devtools [@devtools2021] and pkgdown [@pkgdown2022] to run continuous integration checks on GitHub.  --><!-- These tests can include those of units (do individual functions produce expected output?), documentation (does documentation render correctly?; can all example workflows be executed?) and installation (can the software be successfully deployed on multiple types of operating system?).  -->

<!-- The package lifeycle [@lifecycle2021] provides tools for R developers to consistently deprecate their code. -->

<!-- Guidelines: -->

<!-- - **I1: Model code and data are version controlled** <!-- Was I2 -->

<!-- - **I2: Model code and data use semantic versioning** <!-- Was I1 -->

<!-- - **I3: Continuous integration is used to verify model code updates** -->

<!-- - **I4: Deprecation conventions are used to retire model code and data** -->

>>>>

@@

Model code and data should be clearly documented, potentially with different versions for technical and non-technical users [@Eddy2012]. Consistent use of meaningful naming conventions when authoring code is recommended [@Wilson_2017; @Alarid2019]. Code can be made easier to follow by using the practices of abstraction [@8717448], where only simple, high level commands are routinely exposed to reviewers, and polymorphism [@7181447], where the same command (e.g. "simulate") can be reused to implement different algorithms of the same type. Programs to implement model analyses can be made comprehensible to even non-technical users through the use of literate programming techniques and tools like RMarkdown [@xie2018r] that integrate computer code with plain English descriptions.

@@

### Standards for a reusable MOSHEM
```{r echo = F}
gpl_pc_dbl <- get_license_share("GPL")
```

@@

Storing model code and data in separate files and locations (as opposed to hard coding - embedding data into source code) can make it easier to apply models to different decision contexts and, when necessary, to selectively restrict access to data that are confidential, while disseminating all other model artefacts. Clear distinctions should be made between model modules (code that defines abstract data structures and the algorithms that can be applied to data described by these structures), model datasets (digital information such as parameter values, unit records, etc) and model analyses (code that links model datasets to model modules and specifies the algorithms to apply to the data associated with each module).

The coding practice of encapsulation [@8717448] can be used to help ensure that model modules continue to work as intended even when they are combined [@ready4oop2022]. In some cases, combining modules may mean new versions of modules have to be created to better account for interaction effects. The coding concept of inheritance [@8717448] can be used to efficiently achieve this objective as well as to facilitate selective editing of modules when transferring models to different decision contexts [@ready4oop2022]. Writing algorithms as collections of functions (short, self-contained and reusable software routines that each perform a discrete task) is recommended as good practice for scientific computing [@Wilson_2017]. Functions to implement model algorithms can be associated with module data structures (also known as a class) via a special type of function called a method. Model modules of a similar type or purpose can be efficiently distributed and documented by bundling them as code libraries. It is good practice to make available test or toy data to demonstrate the use of model algorithms [@Wilson_2017].

Statistical models are a common output of health economic evaluations, but they are often not reported in a format that enables others to confidently and reliably re-use them [@Kearns2013]. Open source approaches can help address this by disseminating code artefacts that enable easy and appropriate use of a statistical model to make predictions with new data. However, great care must be exercised when doing so if models are derived from data on human subjects as some software artefacts by default contain a copy of the source dataset. Such dataset copies must therefore be replaced (for example, with synthetic data) and the amended artefact's predictive performance then retested before any public release. Another way to make MOSHEMs easier to use is to develop simple user-interfaces for non-technical users. In the open source R language, such user-interfaces are typically developed with the Shiny package, for which a tutorial aimed at health economists is available [@SmithR2020].

### Standards for an updatable MOSHEM
To avoid MOSHEMs going stale - losing validity and usefulness with time - they should be routinely updated.  Each update of code and data should be uniquely identifiable and retrievable, a goal that can be facilitated by use of version control tools [@Erdemir2020].

Potential users of model code and data should be able to easily identify which version is most appropriate to their needs. Using semantic versioning [@semver20xx] conventions can signal the potential importance of an update to users of model code and data. Also informative is to clearly label the type of code hosted in a repository as either:

- "development" (typically the most comprehensive and up to date source code, but potentially not yet sufficiently well documented and tested for potential users to apply to purposes other than testing);

- "production" (code that has been released with a view to it being applied to its stated purposes and therefore typically required to meet defined documentation and testing standards); or

- "archive" (permanent copies of code at key milestones in its development - potentially useful if seeking to use older versions of code to reproduce analyses).

Continuous integration [@CI2017] tools can help verify that each code update passes multiple quality tests. Finally, using deprecation conventions that take an informative and staged approach to retiring old code and data reduces the risk that model revisions have unintended consequences on third party users.  The package lifeycle [@lifecycle2021] provides tools for R developers to consistently deprecate their code.

### 2023-01-26 Markdown/Manuscript/Child_RMDs/MS_Imp_Toolkit.Rmd
working on discussion

---
title: "MS_Imp_Guidelines"
output: html_document
date: '2022-07-21'
---
<!-- # Framework -->
<!-- The framework we have developed to implement **readyforwhatsnext** specifies standards for OSHEMs and provides tools for meeting those standards. -->

We have developed a toolkit to help streamline the process of developing OSHEMs that meet the TIMELY standards. The toolkit is comprised of online repositories and software.

## Repositories <!-- Existing ref -->
We created a GitHub organisation where all framework software source code is stored, documented, version controlled and continuously integrated [@ready4gh2022]. To store citable archived copies of release copies of our software, we created a Zenodo community [@ready4zen2022]. Finally, to manage datasets for use in models developed with the framework, we created a dedicated collection within the Harvard Dataverse  [@ready4dv2022].

## Software <!-- Novel ref -->

A coding framework for OSHEMs developed in the language R includes standardised approaches to directory structure and naming conventions [@Alarid2019].

As a foundation for implementing the framework, we authored a development version R package that defines a novel syntax and a template class for model modules [@ready42022]. To enable the syntax and module template be applied to modelling projects, we then created five additional development version R packages that provide tools for authoring:

- documented model modules [@ready4class2022];

- documented functions (including methods), written in a consistent house style [@ready4fun2022];

- citable, quality assured R packages [@ready4pack2022];

- model datasets [@ready4use2022];

- model analyses and reports [@ready4show2022].

The six R packages, their primary focus, the TIMELY standards they support and the third-party packages they depend on are summarised in Table \@ref(tab:cpkgs). When used in conjunction with framework repositories, the six packages extend existing R packages provide strong support for implementing 16 of the TIMELY standards. However, the software only weakly supports implementing the standards relating to disseminating statistical models (Y1) and user-interface development (Y2) and does not yet provide any workflow tools to help implement the standards for code coverage (T5) and deprecation conventions (I4). Standards not supported or weakly supported by our software can be met with existing developer tools in R and we plan to progressively integrate these third-party tools with our own in future releases of our software. Another future priority is to submit production versions of each R package for review by and archiving on the Comprehensive R Archive Network (CRAN) [@CRAN2022].

>>>>

---
title: "MS_Imp_Guidelines"
output: html_document
date: '2022-07-21'
---

We developed a toolkit to help us develop and use a MOSHEM that meets the standards listed in Table \@ref(tab:timelygls). The toolkit is comprised of online repositories and software.

## Repositories <!-- Existing ref -->
We created a GitHub organisation (a collection of code repositories) where all our development code is stored and version controlled [@ready4gh2022]. We linked our GitHub organisation to an account we established at codecov [@codecov_2022] to track our code coverage. We configured the repositories in our GitHub organisation to use GitHub actions to support continuous integration.

For hosting documentation websites for the code libraries we create, we enabled GitHub Pages in each library's repository. We also developed a consolidated project documentation website (https://www.ready4-dev.com), using the Hugo framework [@hugo_2023] and Docsy theme [@docsy_2023]. To host that website we established an account with Netlify [@netlify_2023] and linked that service to our GitHub organisation so that the website would automatically update whenever the source code in its GitHub repository was edited.

We also created a Zenodo community [@ready4zen2022] - a collection of permanent, unbiquely identified repositories. We then linked our Zenodo community and GitHub organisation so that every time we create a new "release" in a GitHub repository, a copy of that code is automatically created on Zenodo with a DOI. Finally, to manage model datasets, we created a dedicated collection within the Harvard Dataverse installation [@ready4dv2022].

## Software <!-- Novel ref -->
We created six development version R libraries to help us author model modules, supply them with data and implement reproducible modelling analyses. The six libraries, their primary focus, the standards they support and the third-party packages they depend on are summarised in Table \@ref(tab:cpkgs).

The foundation for our MOSHEM is provided by the ready4 library [@ready42022] that defines a template module (using R's S4 type class) from which all model module data structures will inherit features and a novel syntax for implementing the algorithms attached to model modules. The ready4 library also contains tools for retrieving web based information on model modules, datasets and analysis programs and for partially automating updates to the project website.

Three R libraries are designed to standardise and partially automate workflows for authoring new model modules. The ready4pack library [@ready4pack2022] provides tools for authoring module libraries that are:

- documented (with a website, a manual itemising selected contents and a manual itemising all contents);

- licensed (with GPL-3 by default);

- easily citable (citation information can be retrieved within an R session or from hosting repositories); and

- quality assured (triggering continuous integration workflows, including any unit tests created by module library authors).

The ready4pack library is designed to integrate with our GitHub organisation and depends on the other two module authoring libraries. Methods from the ready4fun library [@ready4fun2022] are used to verify that functions for implementing module algorithms are written in a consistent house style and to use their standardised format to automatically generate basic documentation for each function. Methods from the ready4class [@ready4class2022] library are used to streamline and standardise the authoring of module data-structures. Like ready4fun, the ready4class library uses standardised naming conventions to automatically generate basic documentation for each module data-structure.

The ready4use [@ready4use2022] library contains modules for ingesting model datasets from online repositories (hosted on a Dataverse installation or on GitHub), labelling model datasets and sharing model datasets via online repositories. The ready4show library [@ready4show2022] contains tools to help author analyses programs that are either self-documenting or which trigger the creation of a scientific summary.

When used in conjunction with repositories, the six R libraries provide strong support for implementing 17 framework standards. However, the software only weakly supports implementing the standards relating to safe dissemination of statistical models (R8) and user-interface development (R9) and does not yet provide any workflow tools to help implement the standard for deprecation conventions (U4). Standards not supported or weakly supported by our software can be met with existing developer tools in R.
